\section{SimPoint}
\label{sec:background:simpoint}

When performing experiments via simulation, it is common practice to use excerpts of benchmarks to represent the entire benchmark.
Using excerpts is often a required simplification because simulating an entire benchmark can in many cases take hours or days.
When performing multiple experiments, this delay adds up and can quickly become the factor limiting the size and number of experiments.

A naive approach for extracting excerpts from benchmarks is to specify a number of cycles into the execution where the excerpt starts, in addition to the length of the excerpt.
The reason an excerpt does not usually start at the first executed instruction is because of the startup effect.
When an application starts, it usually performs some initialization tasks before the main routines of the application starts.
During this startup phase, the application may not behave as it will during the remainder of its execution.
As a result, simulating this startup phase might yield unexpected results that are not representative for the benchmark as a whole.
The problem with the naive approach to extracting excerpts is that it is hard to define an offset, and then guarantee that this offset is sufficiently large to skip the startup phase for all benchmarks.

More sophisticated approaches to this problem include tools that analyze the instruction stream of a benchmark as it executes.
Based on this analysis the tool will attempt to locate one or more intervals that based on a selected metric can be considered representative for the entire benchmark execution.
SimPoint\cite{Hamerly2005} is one such tool, based on Pin~\cite{Luk2005}.
Put simply, SimPoint will analyze the benchmark code and divide it into basic blocks.
A basic block is a group of instructions with only a single entry point, the first instruction, and a single exit point, at the last instruction.
It wil then analyze the running program, and each execution interval is classified using a Basic Block Vector (BBV)
The BBV contains an entry for each basic block in the program and a counter indicating how many times the interval executed this block.
After classifying all intervals, SimPoint uses K-means clustering to create K groups of intervals that have similar BBVs.
Within each cluster, the interval closest to the cluster center is selected to represent the cluster. 
The output of SimPoint is the start and length of each cluster's representative interval, along with a weight indicating how much of the entire program execution this cluster represents.

Previous work by Hamery~et~al.~\cite{Hamerly2004} has compared the simulation results of SimPoint selected intervals and intervals selected by the naive approach to the results of simulating the entire benchmark.
When using SimPoint selected intervals each interval is simulated independently, and a weighted sum of simulation results is created using the generated weights.
Hamerly~et~al. show that using SimPoint selected intervals can improve the accuracy of the simulation results. 
They also show that the value of K, the number of intervals returned, and the length of the interval effect the accuracy of the solution. 

\todo{Extend the last sentence with some actual results from the Hamerly et al. paper}